---
title: "Testing and correcting for publication bias"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, error=FALSE, message=FALSE)
```

```{r libraries, echo=FALSE}
library(RMySQL)
library(metafor)
library(tidyr)
library(dplyr)
library(stringr)
library(purrr)
library(ggplot2)
```

This document documents the analysis of the project "testing and correcting for publication bias".
The goal of the project is to test several methods of testing of and correction for publication bias both on simulated data and on true data.
The most basic idea is to emulate Kvarven et al (2019) NHB paper using datasets from replications in psychology as a benchmark and confronting them with the results of meta-analysis published before the replications took place. 
This is still needed since the results in Kvarven are still open to discussion, are not fully reproducible (with a full-on reproducible program) and do not cover all methods.
The hope is also to set up a benchmark for accumulating evidence on the extent of publication bias and on the performances of methods testing and correcting for it.

# Empirical assessment of funnel plots

The first order of business is to try to build a very quick bird's eye view of the extent and shape of selection bias in the 15 datasets used by Kvarven.
They have all been downloaded from the OSC repository of the Kvarven paper into a separate MySQL server, so that all calls to the datasets are normalized.
Let's first download the datasets. 

```{r libraries, echo=FALSE}
## On Mac, this is the way to read files
source(here::here("idsql.R"))
kvarven <- dbConnect(MySQL(), dbname="Kvarven",
                     group='LaLonde_amazon',
                     user=myid, password=mypass, host=myhost)
# list of datasets
names.datasets <- dbGetQuery(kvarven,"SELECT `TABLE_NAME` FROM `information_schema`.`COLUMNS` WHERE (`TABLE_SCHEMA` = 'Kvarven');") %>%
                    pull(TABLE_NAME) %>%
                    unique()
# dowload data
fromSQL <- names.datasets %>% map(~dbReadTable(kvarven, .)) %>% set_names(names.datasets) # returns a list with as many elements as datasets for this paper_id
```

Now, for each dataset, we want to plot the funnel plot along with the replication-based estimate of the true effect size.
One way to do that is to regroup all datasets into one and to use the **facetwrap** function of **ggplot**.
But first, we have to define a new variable taking as value the name of the dataset in each dataset, which will be our grouping variable in the facets.
Let's see if I can do that.

```{r FullData}
Studies <- fromSQL[2:16]
# function generating a new column in a dataframe whose value is unique and in a vector
NewColumnFun <- function(name,data){
  data <- data %>%
            mutate(
              Study = name
            )
  return(data)
}

# New list of datasets
Studies <- map2(names.datasets[-1],Studies,NewColumnFun)
names(Studies) <- names(fromSQL[-1])

# one dataset with all studies
DataFull <- Studies %>%
              bind_rows()
```

I also have to prepare the aggregate data to extract the name of the meta-analysis and the original study so that it has the same and has a correspondance.

```{r AggregateData}
Aggregate <- fromSQL[[1]] %>%
              mutate(
                Study = str_split_fixed(metaanalysis," ",n=2),
                Original = str_split_fixed(original," ",n=2)
              )
Aggregate[["Study"]] <- Aggregate[["Study"]][,1]
Aggregate[["Original"]] <- Aggregate[["Original"]][,1]

# original study name in the full dataset
DataFull <- DataFull %>%
              left_join(select(Aggregate,Study,Original),by="Study")
```

Let's now plot the data.

```{r FunnelPlotsFullData}
ggplot(DataFull,aes(x=sed,y=d)) +
  geom_point()+
  geom_hline(data=Aggregate,aes(yintercept=replication_s),color='red',linetype='dashed') +
  geom_hline(data=Aggregate,aes(yintercept=meta_s),color='green',linetype='dashed') +
  geom_hline(data=Aggregate,aes(yintercept=effecto),color='blue',linetype='dotted') +
  coord_cartesian(ylim=c(-1,2))+
  facet_wrap(~Original)+
  theme_bw()
```
